% !TEX root =  ../Dissertation.tex

\chapter{Legal, Social, Ethical and Professional Issues}


The project uses several open-source libraries, such as NetworkX and sexpdata.
It is important to comply with the licenses of these libraries, such as MIT and
BSD to avoid legal issues. Without compliance, this tool can't be made publicly
available. This project does not handle sensitive user data, any future
additions that might involve user data must comply with data protection
regulations.

The project has the potential for a positive impact on developer productivity by
giving tools to understand large codebases quickly and reducing downtime
during refactoring. This tool needs to be accessible and
inclusive to a wide range of users. Since the tool runs on a terminal the user
can choose the interface that suits them best. It is also important to provide
documentation and helpful error messages to help users unfamiliar with Python
to use this tool.

Transparency and accountability are critical for a project that will analyze
the personal and professional projects of developers. The tool must give
accurate results when querying definitions and creating compilation order.
Misleading output can introduce errors during development which is unethical
and counterproductive.

This project must follow the professional standards set by the BCS Code of Conduct.
These standards ensure that software is reliable, secure, and ethical. The tool
will evolve and it is critical for it to be well-tested, to avoid
introducing errors.

% Address the legal, social, ethical, and professional issues associated with
% your project. Find something to say for each. For example, listening for legal,
% BCS Code of Conduct for professional, etc.
%
% \begin{itemize}
% \item Legal consider licensing of third-party libraries and how they can be
%     used. Data protection isn't an issue as it all runs locally.
% \item Social: the tool provides documentation so it can be used by a variety of
%     developers. Tool runs in any terminal so it can be customized to fit any
%     accessible need from text. 
% \item Ethical: the source code for the tool is fully open for transparency and
%     collaboration. 
% \item Professional: Adherence to the BCS Code of Conduct.
% \end{itemize}

